---
title: "The Capacities Of Different Homeless Shelter Sectors"
author: "Mahmoud Elsheikh"
date: '2020-09-24T21:13:14-05:00'
categories: R
tags:
- R Markdown
- plot
- regression
---
---
title: "The Capacities Of Different Homeless Shelter Sectors"
output: html_document
---

```{r}
library(devtools)
library(decon)
library(ggplot2)
library(tibble)
library(ggthemes)
library(RColorBrewer)

```


# Abstract
This is a statistical analysis of the capacity of homeless shelters in the City of Toronto. This analysis aims to find the different capacities in homeless shelter sectors. This will alow for a better understanding of which sectors have excess capacities and which are struggling to meet the demand.

# Introduction
Homelesness is a serious issue around the country and Toronto is not exempt from that. On any given night there are 9,200 homeless people in Toronto^1 in need of a place to stay. While homeless shelter's take in a significant proportion of the homeless, they are unable to cater to evrybody which you will see in the data below. The dataset I selected to use was the Daily Shelter Occupancy Data. I found it very interesting to find out which shelter sector's are more or less catered for through analysing their excess capacity. I found this very interesting as I have made use of the violin geom in order to add the a third dimension to the graph of the capacity density's.



The dataset being used is the daily shelter occupancy data obtained from the Toronto Open Data Portal providing a listing of all active shelters in the City of Toronto area in 2020. This dataset contains data from each shelter ranging from 01-01-2020 to 21-09-2020

The graph below displays the available capacity for different sectors of shelters every day from 01-01-2020 to 21-09-2020. The sectors refer to the gender or age group  the shelter caters to. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

ggplot(data = Daily.shelter.occupancy_current., aes(x = SECTOR, y = CAPACITY, fill = SECTOR)) + geom_point(aes(fill = SECTOR)) + geom_violin(scale="area") + theme_solarized()
```


## The data is very interesting as it shows us each sector shelter's ability to accept people. This data also gives us insight into the 


# Citations

1. https://www.fredvictor.org/facts-about-homelessness-in-toronto/#:~:text=Facts%20about%20Homelessness%20in%20Toronto,homeless%20on%20any%20given%20night.

  H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New
  York, 2016.

A BibTeX entry for LaTeX users is

  @Book{,
    author = {Hadley Wickham},
    title = {ggplot2: Elegant Graphics for Data Analysis},
    publisher = {Springer-Verlag New York},
    year = {2016},
    isbn = {978-3-319-24277-4},
    url = {https://ggplot2.tidyverse.org},
  }


  Jeffrey B. Arnold (2019). ggthemes: Extra Themes, Scales and Geoms for 'ggplot2'.
  R package version 4.2.0. https://CRAN.R-project.org/package=ggthemes

A BibTeX entry for LaTeX users is

  @Manual{,
    title = {ggthemes: Extra Themes, Scales and Geoms for 'ggplot2'},
    author = {Jeffrey B. Arnold},
    year = {2019},
    note = {R package version 4.2.0},
    url = {https://CRAN.R-project.org/package=ggthemes},
  }



Hadley Wickham, Jim Hester and Winston Chang (2020). devtools: Tools to Make
  Developing R Packages Easier. R package version 2.3.2.
  https://CRAN.R-project.org/package=devtools

A BibTeX entry for LaTeX users is

  @Manual{,
    title = {devtools: Tools to Make Developing R Packages Easier},
    author = {Hadley Wickham and Jim Hester and Winston Chang},
    year = {2020},
    note = {R package version 2.3.2},
    url = {https://CRAN.R-project.org/package=devtools},
  }


