---
title: "The Capacities Of Different Homeless Shelter Sectors"
author: "Mahmoud Elsheikh"
date: 2015-07-23T21:13:14-05:00
categories: ["R"]
bibliography: "references.bib"
tags: ["R Markdown", "plot", "regression"]
---
---
title: "The Capacities Of Different Homeless Shelter Sectors"
output: html_document
---

```{r}
library(devtools)
library(decon)
library(ggplot2)
library(tibble)
library(ggthemes)
library(RColorBrewer)

```


# Introduction
## The dataset I selected to use was the Daily Shelter Occupancy Data. I found it very interesting to find out which shelter sector's are more or less catered for as far as having excess capacity. I found this very interesting as I have made use of the violin geom in order to add the demension of the capacity density's 



## The dataset being used is the daily shelter occupancy data obtained from the Toronto Open Data Portal providing a listing of all active shelters in the City of Toronto area in 2020. This dataset contains data from each shelter ranging from the 
The dataset contains data from 01-01-2020 to 21-09-2020

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

ggplot(data = Daily.shelter.occupancy, aes(x = SECTOR, y = CAPACITY, fill = SECTOR)) + geom_point(aes(fill = SECTOR)) + geom_violin(scale="area") + theme_solarized()


```

## The graph below displays the available capacity for different sectors of shelters every day from 01-01-2020 to 21-09-2020. The sectors refer to the gender or age group  the shelter caters to. 
## The data is very interesting as it shows us each sector shelter's ability to accept people. This data also gives us insight into the 


